---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# whitewater

<!-- badges: start -->
[![R-CMD-check](https://github.com/joshualerickson/whitewater/workflows/R-CMD-check/badge.svg)](https://github.com/joshualerickson/whitewater/actions)[![codecov](https://codecov.io/gh/joshualerickson/whitewater/branch/main/graph/badge.svg)](https://app.codecov.io/gh/joshualerickson/whitewater)
<!-- badges: end -->


The goal of whitewater is to provide sequential and parallel processing for USGS stations in a tidy-style format. 

## Installation

You can install the development version of whitewater from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("joshualerickson/whitewater")
```

## Example

This is a basic example which shows you how to solve a common problem: get minimum flows per water year for multiple sites using parallel processing. Please see [furrr](https://cran.r-project.org/web/packages/furrr/index.html) and [future](https://cran.r-project.org/web/packages/future/index.html) for more details on parallel processing methods.  

### Running in parallel  

```{r}
library(whitewater)
library(ggplot2)
library(future)

#need to call future::plan()
plan(multisession(workers = workersAvailable()-1))

```


```{r example}

yaak_min <- ww_wyUSGS(sites = '12304500')

ggplot(yaak_min, aes(wy, Flow_min)) +
  geom_point() + 
  geom_line() +
  theme_bw()
```
### Or minimum/maximum water temperature

```{r}

withlacoochee_temp <- ww_wyUSGS(sites="02319394",
                          parameter_cd = c("00010"))

ggplot(withlacoochee_temp, aes(wy, Wtemp_min)) +
  geom_point() + 
  geom_line() +
  theme_bw()

ggplot(withlacoochee_temp, aes(wy, Wtemp_max)) +
  geom_point() + 
  geom_line() +
  theme_bw()
```

