---
title: "Get Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{get_dv}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev.args = list(png = list(type = "cairo"))
)

ggplot2::theme_set(ggplot2::theme_bw())
```

## Intro  

The reason for whitewater has always been to get USGS water data with parallel processing. However, there's also been the idea of tidying up the returned data. This vignette will go into using some of the functionality in whitewater.  

## ww_*()'s  

All of the functions in whitewater begin with a prefix `ww`. This is to help the user search for functions quickly using [Rstudios code completion](https://support.rstudio.com/hc/en-us/articles/205273297-Code-Completion-in-the-RStudio-IDE) as well as simply remember! 

## Flow  

The main flow in whitewater is to start with `ww_dvUSGS()` and then the user can use that object for further filtering within functions, e.g. `ww_wyUSGS()`, `ww_wymUSGS()`, `ww_monthUSGS()`, etc. We'll start with just getting daily values of water data and then use that object to supply to the others. Of note, you can specify the water year `wy_month`, which is the beginning month for that water year.  

```{r setup}
library(whitewater)
library(dplyr)
library(ggplot2)
library(tidyr)

yaak_dv <- ww_dvUSGS(sites = '12304500',
                     parameter_cd = '00060',
                     wy_month = 10)
```

This is just a wrapper around `dataRetrieval::readNWISdv()` except that we've added some attributes along the way; drainage area, lat and long, altitude, observations per water year `obs_per_wy`, observations per month `obs_per_month`, day of year `doy`, day of water year `wy_doy` as well as parsing out the date to `month`, `day`, `year` and `wy`. As you'll see below, this helps with EDA and data wrangling.  

```{r, fig.align='center', fig.height=6, fig.width=8}
yaak_dv %>% 
  ggplot(aes(wy_doy, Flow)) + 
  geom_line(aes(group = wy, color = wy)) +
  scale_color_gradientn(colors = hcl.colors(n = 11, palette = 'Zissou1'))
```


### Water Year  

Now if we just want per water year we can call the function `ww_wyUSGS()`. This function will call the `dataRetrieval::readNWISPeak()` function in the background if `parameter_cd = '00060'` is used.  

```{r}
yaak_wy <- yaak_dv %>% ww_wyUSGS()
```

We now have a `tibble()` with a bunch of summary statistics; mean, maximum, median, standard deviation and some normalization methods (drainage area, scaled by log and standard deviation) per water year. In addition, the function also performs some counting methods as mentioned above so that we can filter out water years that don't have all (or close to all) the observations. Below you can see that we now have only water years with greater than 360 observations.  

```{r}

yaak_wy %>% filter(obs_per_wy > 360) %>% select(Station, wy, obs_per_wy) %>% count(obs_per_wy)
yaak_wy_filter <- yaak_wy %>% filter(obs_per_wy > 360) 
```

This will make it nice if you wanted to see the dependence between water year minimum and sum flow. If we didn't filter we'd be skewing observations since it doesn't have all of the data!  

### Water Year and Month  

Another nice tidying process is the summary statistics for all the water years and months. By calling `ww_wymUSGS()` we can can get this.  

```{r}

yaak_wym <- yaak_dv %>% ww_wymUSGS()

```
Again, since we have `doy`, `obs_per_month` and `wym_count` we can filter out water years with less than 12 months and then plot to see monthly statistics over time.  

```{r, fig.align='center', fig.height=6, fig.width=8}
yaak_wym %>% 
  filter(wym_count == 12, 
         obs_per_month > 27) %>% 
  pivot_longer(cols = contains("Flow")) %>% 
  ggplot(aes(doy, value, group = wy, color = wy)) + 
  geom_line() + 
  facet_wrap(~name, scale = 'free')
```

### Months  

Now if we want only month statistics over time we just use the `ww_monthUSGS()`. This is essentially just grouping by month and then running the summary statistics of the daily values.  

```{r}

yaak_month <- ww_monthUSGS(yaak_dv)
```
### Stats  

Sometimes you just want to compare the current flow (water year or past week, whatever you want) to historical flows. The `ww_statsUSGS()` function does this for you! It takes the historical values for your parameter (flow in this example) and returns percentiles (`dataRetrieval::readNWISstat()`) but also combines the current values.  

```{r, fig.align='center', fig.height=6, fig.width=8}

yaak_daily_report <- yaak_dv %>% ww_statsUSGS(temporalFilter = 'daily',
                                  days = 365)
yaak_daily_report %>% 
  pivot_longer(c('Flow', 'p25_va', 'p75_va', 'max_va', 'min_va')) %>% 
  ggplot() + 
  geom_line(aes(Date, value, color = name, alpha = name %in% c('p25_va', 'p75_va', 'max_va', 'min_va'))) + 
  scale_color_manual(values = c('black', 'blue', 'red', 'yellow', 'forestgreen')) +
  scale_alpha_manual(values = c(1,.5), guide = 'none') +
  labs(y = 'Discharge (cfs)', color = '', title = 'Comparing current flow to Max, Min, 25th and 75th percentiles') + 
  theme_bw()
```

