---
title: "stats"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{stats}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
   fig.align='center',
  fig.height=6, 
  fig.width=8,
  dev.args = list(png = list(type = "cairo"))
)

ggplot2::theme_set(ggplot2::theme_bw())
```

```{r setup}
library(whitewater)
library(ggplot2)
library(dplyr)
library(tidyr)
library(sf)
library(future)
library(ggfx)
library(geomtextpath)
```


## Intro  

Sometimes you just want to compare the current flow (water year or past week, whatever you want) to historical flows (figure below). The `ww_statsUSGS()` function does this for you! It takes the historical values for your parameter (flow in this example) and returns percentiles for site site but also combines the current values and current conditions if using parameter code '00060' from the [USGS dashboard](https://dashboard.waterdata.usgs.gov/app/nwd/?aoi=default). In this vignette we'll go over the different options/arguments in the `ww_statsUSGS()`.    

```{r}

yaak_dv <- ww_dvUSGS(sites = '12304500', parameter_cd = '00060')

yaak_daily_report <- yaak_dv %>% ww_statsUSGS(temporalFilter = 'daily',
                                  days = 365)
yaak_daily_report %>% 
  pivot_longer(c('Flow', 'p25_va', 'p75_va', 'max_va', 'min_va')) %>% 
  ggplot() + 
  with_outer_glow(geom_line(aes(Date, value, color = name, alpha = name %in% c('p25_va', 'p75_va', 'max_va', 'min_va'))), sigma = .5) + 
  scale_color_manual(values = c('black', 'blue', 'red', 'yellow', 'forestgreen')) +
  scale_alpha_manual(values = c(1,.5), guide = 'none') +
  labs(y = 'Discharge (cfs)', color = '', title = 'Comparing current flow to Max, Min, 25th and 75th percentiles') + 
  theme_bw()
```


## ww_statsUSGS()  

The `ww_statsUSGS()` is essentially a wrapper around `dataRetrieval::readNWISstat()` but in addition will take the parameter values ('Flow', 'Wtemp', etc) calculated from the `ww_floorIVUSGS()`
function and add to the data as well. Thus, the instantaneous values will look different than the daily mean values, as it should. Just like dataRetrieval, there are three different flavors of temporal windows that we can choose from: daily, monthly and yearly.  

### daily  

The `temporalFilter` argument is used to generate the window of percentiles. Like in the example below we just want to get the last 30 days of daily stats so we'll put `temporalFilter = 'daily'` and `days = 30`. The `days` arguments is only used for `daily` but will go back the amount of days you input.  

```{r}
yaak_daily_report_30 <- yaak_dv %>% ww_statsUSGS(temporalFilter = 'daily',
                                  days = 30)
yaak_daily_report_30
```

```{r, echo=F}
yaak_daily_report_30 %>% 
  pivot_longer(c('Flow', 'p25_va', 'p75_va', 'max_va', 'min_va')) %>% 
  ggplot() + 
  with_outer_glow(geom_line(aes(Date, value, color = name, alpha = name %in% c('p25_va', 'p75_va', 'max_va', 'min_va')))) + 
  scale_color_manual(values = c('black', 'blue', 'red', 'yellow', 'forestgreen')) +
  scale_alpha_manual(values = c(1,.5), guide = 'none') +
  labs(y = 'Discharge (cfs)', color = '', title = 'Comparing current flow to Max, Min, 25th and 75th percentiles') + 
  theme_bw()
```

If we want to compare multiple sites that fine too! In addition, you don't have to pipe a `ww_dvUSGS()` object into the function either. You can just input the sites you want; however, I always recommend using `ww_dvUSGS()` because you'll likely come back to the daily values anyway but in the example below we'll just call with inputting the sites into the `sites` argument. If you're going to use the `sites` argument you'll also need to specify the `parameter_cd`!   

```{r}

multiple_sites <- ww_statsUSGS(sites = c('12304500', '14159200'),
                               parameter_cd = '00060',
                               temporalFilter = 'daily', 
                               days = 30)
  

```
```{r, echo = F}

multiple_sites %>% 
  rename(`Current Flow` = 'Flow', `25%` = 'p25_va', `75%` = 'p75_va', Max = 'max_va', Min = 'min_va') %>% 
  pivot_longer(c(`Current Flow`,`25%`, `75%`, 'Max', 'Min')) %>% 
  mutate(name = factor(name, levels = c('Current Flow', 'Max', 'Min', '25%', '75%'))) %>% 
  ggplot() +
  labs(y = 'Discharge (cfs)', color = '', title = 'Comparing current flow to Max, Min, 25th and 75th percentiles') + 
  with_outer_glow(geomtextpath::geom_textpath(aes(Date, value, color = name, label = name, alpha = name %in% c('Current Flow','25%', '75%', 'Max', 'Min')),text_smoothing = 30,hjust = 1, show.legend = F), sigma = .5)+ 
  scale_color_manual(values = c('black', 'blue', 'red', 'yellow', 'forestgreen')) +
  scale_alpha_manual(values = c(1,.5), guide = 'none')  + 
  facet_wrap(~Station, scales = 'free')
```


What's nice about this is you can run in parallel using the argument `parallel = TRUE` (see [Parallel](https://joshualerickson.github.io/whitewater/articles/parallel.html) vignette for more details on parallel processing). We'll use the `ww_current_conditions()` to get the current streamflow status for the lower 48 that has a description of `All-time low for this day` and then we'll use these sites to view the last 30 days!  

```{r, warning=F, message=F, error=F}

current_condition <- ww_current_conditions() %>% filter(StatisticsStatusDescription %in% c('All-time low for this day'))

plan(multisession(workers = 10))

atl_sites <- ww_statsUSGS(sites = unique(current_condition$SiteNumber),
                             parameter_cd = '00060', 
                             temporalFilter = 'daily',
                             days = 30,
                             parallel = TRUE,
                             verbose = FALSE)
```


```{r, echo=F, warning=F, message=F, error=F}

ranking_flow <- atl_sites %>% left_join(current_condition  %>% 
                                          filter(!TimeZoneCode %in% c('AKDT', 'AST','HST')) %>% 
                                             select(site_no = "SiteNumber",
                                                    StatisticsStatusDescription,
                                                    StatisticStatusCode,
                                                    StatisticsStatusColorFill,
                                                    StatisticsStatusColorStroke,
                                                    Latitude,
                                                    Longitude))

mind <- min(atl_sites$Date, na.rm = T)
maxd <- max(atl_sites$Date, na.rm = T)



p1 <- ranking_flow %>% 
  mutate(Flow = if_else(Flow == "NaN", NA_real_, Flow)) %>% 
  filter(!is.na(Flow), !is.na(StatisticsStatusDescription)) %>% 
  ggplot(aes(Date)) + 
  with_outer_glow(geom_line(aes(y = Flow), color = '#FF0000', show.legend = F)) +
  with_inner_glow(geom_ribbon(aes(ymin = p25_va, ymax = p75_va), fill = 'grey75', alpha = 0.2)) + 
  labs(y = 'Discharge (cfs)',
       color = '',
       title = 'Comparing current flow to 25th and 75th percentiles',
       subtitle = paste0('Dates: ', mind,' to ', maxd)) +
  facet_wrap(~site_no, scales = 'free') +
  theme_void() +
  theme(text=element_text(size=10,  family="Comic Sans MS"),
        strip.text = element_blank(),
        axis.title.x = element_text(),
        axis.title.y = element_text(angle = 90,margin = unit(c(0, 3, 0, 0), 'mm')),
        plot.subtitle = element_text(margin = unit(c(0, 0, 3, 0), 'mm')),
        plot.title = element_text(margin = unit(c(0, 0, 3, 0), 'mm')))
p1
```


```{r, echo = F, warning=F, message=F, error=F}

p2 <- ranking_flow %>% 
  group_by(site_no) %>% 
  slice(n = 1) %>% 
  filter(!is.na(Longitude)) %>% 
  ungroup() %>% 
  st_as_sf(coords = c('Longitude', 'Latitude'), crs = 4326) %>% filter(!is.na(StatisticsStatusDescription)) %>% 
  ggplot() + 
  with_inner_glow(borders('state',fill = 'white'),sigma = 4) + 
geom_sf(color = '#FF0000', aes(color = StatisticsStatusDescription,
                               fill = StatisticsStatusDescription),
        stroke = .25, size = 1.25, shape = 21, show.legend = F) +
  scale_fill_manual(values = "#FF0000") + 
  scale_color_manual(values = "#990000") +
  theme_void() +
  labs(
       title = 'USGS Sites with Current Streamflow Status at "All-time low"',
       subtitle = paste0('Date: ', maxd))+
  theme(text=element_text(size=10,  family="Comic Sans MS"))
p2
```

### monthly and yearly   

Now we can looked into the other temporal windows like `monthly` and `annual`. These work the same as daily but now we just need to change the argument.  

```{r}

yaak_monthly_stats <- ww_statsUSGS(yaak_dv, temporalFilter = 'monthly')

yaak_monthly_stats %>% 
  filter(month == 8) %>% 
  mutate(diff = mean_value-p50_va,
         Status = if_else(diff > 0, "Above normal", "Below normal")) %>% 
  ggplot(aes(year_nu, diff, fill = Status)) + 
  with_outer_glow(geom_col(), sigma = 20) +
  scale_fill_manual(values = c("#009900","#996600")) +
  labs(y = 'Discharge (cfs)', x = 'Year', color = '', title = 'Comparing monthly mean flow in August to 50% percentile') + 
  theme_dark()
```


This looks interesting! What about the sites that have all-time low for today? Well, just like above we change the `temporalFilter` argument to `monthly` instead of `daily` and then use those site numbers to get back a tibble with mean monthly flows and statistics for each month per year per site.  

```{r, echo = F, warning=F, message=F, error=F}

atl_sites_monthly <- ww_statsUSGS(sites = unique(current_condition$SiteNumber),
                             parameter_cd = '00060', 
                             temporalFilter = 'monthly',
                             parallel = TRUE,
                             verbose = FALSE)
```

We can then look at a few of the sites in a heatmap. 

```{r, echo = F, warning=F, message=F, error=F}

monthly_stats <- atl_sites_monthly %>% add_count(site_no,year_nu) %>% filter(n == 12)

atl_sites_monthly %>% 
  group_by(site_no) %>%   
  add_count() %>% 
  filter(n > 950) %>% 
  nest() %>% 
  .[1:4,] %>% 
  unnest() %>% 
  ungroup() %>% 
  ggplot() + 
  with_inner_glow(geom_tile(aes(month, year_nu, fill = StatisticsStatusDescription))) +
  scale_x_continuous(breaks = scales::pretty_breaks()) + 
  scale_fill_manual('Streamflow: status',
                    values = c(
                      "All-time low for this month" = "#FF0000",
                      "Much below normal" = "#BB2222",
                      "Below normal" = "#FFAA00",
                      "Normal" = "#00ff00",
                      "Above normal" = "#44dddd",
                      "Much above normal" = "#00ffff",
                      "All-time high for this month" = "#000055")) + 
  labs(title = "Heatmap of Monthly Mean Streamflow", x = 'Month', y = 'Year')+
  facet_wrap(~Station, scales = 'free')+
  theme(text=element_text(size=10,  family="Comic Sans MS"),
        strip.text = element_text(size = 5),
        axis.title.x = element_text(),
        axis.title.y = element_text(angle = 90,margin = unit(c(0, 3, 0, 0), 'mm')),
        plot.subtitle = element_text(margin = unit(c(0, 0, 3, 0), 'mm')),
        plot.title = element_text(margin = unit(c(0, 0, 3, 0), 'mm')))
```

**Or, back to the Yaak, MT station.**  

```{r, echo = F, warning=F, message=F, error=F}
yaak_monthly_stats %>% 
ggplot() + 
  with_inner_glow(geom_tile(aes(month, year_nu, fill = StatisticsStatusDescription))) +
  scale_x_continuous(breaks = scales::pretty_breaks()) + 
  scale_fill_manual('Streamflow: status',
                    values = c(
                      "All-time low for this month" = "#FF0000",
                      "Much below normal" = "#BB2222",
                      "Below normal" = "#FFAA00",
                      "Normal" = "#00ff00",
                      "Above normal" = "#44dddd",
                      "Much above normal" = "#00ffff",
                      "All-time high for this month" = "#000055")) + 
  labs(title = "Heatmap of Monthly Mean Streamflow", x = 'Month', y = 'Year')+
  facet_wrap(~Station, scales = 'free')+
  theme(text=element_text(size=10,  family="Comic Sans MS"),
        strip.text = element_text(size = 5),
        axis.title.x = element_text(),
        axis.title.y = element_text(angle = 90,margin = unit(c(0, 3, 0, 0), 'mm')),
        plot.subtitle = element_text(margin = unit(c(0, 0, 3, 0), 'mm')),
        plot.title = element_text(margin = unit(c(0, 0, 3, 0), 'mm')))
```

Same as monthly we can do this for `yearly`.  

```{r}

yaak_yearly_stats <- ww_statsUSGS(yaak_dv, temporalFilter = 'yearly')
```


Now we can look at the mean value for the year against the 25th and 75th percentiles.  

```{r, echo = F, warning=F, message=F, error=F}
yaak_yearly_stats %>% 
ggplot(aes(year)) + 
  with_outer_glow(geom_line(aes(y = mean_value), color = 'black', show.legend = F),sigma = 2, expand = .1) +
  with_outer_glow(geom_ribbon(aes(ymin = p25_va, ymax = p75_va), fill = 'grey75', alpha = 0.2), sigma = 10) + 
  labs(y = 'Discharge (cfs)',
       color = '',
       title = 'Comparing yearly mean flow to 25th and 75th percentiles') +
  facet_wrap(~Station, scales = 'free') +
  theme(text=element_text(size=10,  family="Comic Sans MS"),
        strip.text = element_blank(),
        axis.title.x = element_text(),
        axis.title.y = element_text(angle = 90,margin = unit(c(0, 3, 0, 0), 'mm')),
        plot.subtitle = element_text(margin = unit(c(0, 0, 3, 0), 'mm')),
        plot.title = element_text(margin = unit(c(0, 0, 3, 0), 'mm')))
```

### Wrap-up  

Hopefully this gives you an idea of how to use whitewater with the `ww_statsUSGS()` function? If you have any questions or issues please submit a issue on [github](https://github.com/joshualerickson/whitewater/issues). Thanks!  
